\maketitle
\bibliographystyle{nar}
\label{basics}
In the following guide we'll show  a step by step introduction to 3DNA
from installation of the software to practical examples.

\section{Installation}
To install 3DNA you will first have to download the software from:\\

\url{http://adna.rutgers.edu}\\

\noindent The webpage will ask you to register before allowing you to
download the program.\\
Once   registered    you'll   be    able   to   download    the   file
\textbf{Linux\_X3DNA\_v2.0.tar.gz} if you  selected the Linux version.\\
To uncompress issue the following command in a terminal
\begin{Verbatim}
tar -xvzf Linux_X3DNA_v2.0.tar.gz
\end{Verbatim}
This will create a new folder named \textbf{X3DNA} which you will have
to  include  in your  environment  variables  by  changing your  shell
environment  configuration   file  named  ``.bashrc''   in  bash,  and
``.chsrc'' in csh and tcsh.\\ 
If you are using .bashrc the following lines will need to be added:
\begin{Verbatim}
export X3DNA=/home/yourusername/X3DNA
export PATH=$PATH:$X3DNA/bin
\end{Verbatim}
A detailed explanation  of how to configure X3DNA can  be found in the
pdf file \textbf{x3dna\_v1.5.pdf} found in the X3DNA/doc folder.

\section{fiber}
The \textrm{fiber} program included in  the 3DNA package allows you to
easily  generate  protein data  bank  (pdb)  files of  fiber
models of common nucleic acid conformations, for example A-DNA, B-DNA,
and Z-DNA.  To get a list of  all possible fiber models  which you can
reconstruct using \textrm{fiber}, type:
\begin{Verbatim}
fiber -l
\end{Verbatim}
If you type \textrm{fiber} in the terminal you  will get a usage
message.\\
Stop and read the usage message carefully.\\
Now, you  know that the crystal structures of A-RNA and B-DNA must
have a total of 11 and 10 residues per turn respectively, so, go ahead
and use  \textrm{fiber} to make  an A-RNA with  12 residues and  a B-DNA
with  11 residues.  Using a  molecular visualization  program  such as
pymol, vmd,  or chimera check  the generated structure  to corroborate
that you have the correct number of residues per turn.

\section{find\_pair}
The \textrm{find\_pair} program of 3DNA  allows you to find pairs in a
given pdb file.
To read the usage message open a terminal and type:
\begin{Verbatim}
find_pair
\end{Verbatim} 
Now go  to the protein data bank website  at \url{http://www.pdb.org}
and download the  structure with PDB\_ID: 1ehz. This  is the structure
for yeast phenylalanine tRNA.
Run it through \textrm{find\_pair}:
\begin{Verbatim}
find_pair 1ehz.pdb 1ehz.inp
\end{Verbatim}
You will  get a file named  1ehz.inp. Go ahead and  open the resulting
file in  a text editor and  you will see  that the \textrm{find\_pair}
program has found the  canonical Watson-Crick base-pairs G$\cdot$C and
A$\cdot$U, as  well as non-canonical base-pairs such  as G$\cdot$U and
A$\cdot$G.

\section{analyze}
The \textrm{analyze}  program of the 3DNA software  package allows you
to compute the base-pair and  base-pair step parameters associated with a
given nucleic  acid structure among  other analysis features  such as
computing the overlap areas  between stacked nucleic acid bases, sugar
conformations, and sugar-phosphate backbone torsion angles.\\
A requirement of the \textrm{analyze} program is that you have to run the
\textrm{find\_pair} program beforehand.\\
Go ahead and read the usage message of the analyze program by typing:
\begin{Verbatim}
analyze --help
\end{Verbatim}
in your terminal.\\

\noindent Now run the analyze  program on the file previously produced
using the \textrm{find\_pair} program:
\begin{Verbatim}
analyze 1ehz.inp
\end{Verbatim}
Notice that  you can run the  \textrm{find\_pair} and \textrm{analyze}
program in just one line in the terminal by issuing the command:
\begin{Verbatim}
find_pair 1ehz.pdb stdout | analyze
\end{Verbatim}
This  command   will  produce  a  wealth  of   files  with  structural
information, but  the main one  will just have  the name of  the input
file followed  by the .out extension.  That is, you should  now have a
1ehz.out file in your folder. Go ahead and take a look at the 1ehz.out
file in your favorite text editor.

\section{rebuild}
The very useful 3DNA  program called  \textrm{rebuild} can  do just
that, rebuild nucleic acid structures. It uses as input either a set of base
step, or base-pair step parameters, or a set of helical parameters to
create a pdb file.
To  test  the  rebuild   program  first  run  \textrm{find\_pair}  and
\textrm{analyze} the A-RNA structure you previously created:
\begin{Verbatim}
find_pair A-RNA.pdb stdout | analyze
\end{Verbatim}
Issuing  the previous command must  have created  a new  file called
\textit{bp\_step.par}.
Open this file in a text editor and change the Twist values from 31.5 degrees
to a smaller value, say 25.0 degrees, and save the new file with the name
\textit{undertwisted.par}.\\
Rebuild the undertwisted RNA by typing the following command in your good-ole
terminal:
\begin{Verbatim}
rebuild -atomic undertwisted.par undertwisted.pdb
\end{Verbatim}
This   will  generate   a  pdb   structure  with   no  sugar-phosphate
backbone. To add a sugar-phosphate backbone you will need to issue the
\textrm{cp\_std}  command  before   you  perform  the  \textrm{rebuild}
command, in the following way:
\begin{Verbatim}
cp_std RNA
rebuild -atomic undertwisted.par undertwisted.pdb
\end{Verbatim}
Now  rebuild   a  pdb  file  using  the   unmmodified  base-pair  step
parameters,  that  is,  using  the  \textit{bp\_step.par} file,  and  open  both
structures in  a molecular visualization program and  confirm that one
helix is undertwisted with respect to the original.

\section{Exercises}
1. Run  \textrm{find\_pair} and analyze on your A-RNA structure  and modify the
resulting   \textit{bp\_step.par}  file  so   that  the   sequence  is
AAAGGGUUUCCC. Use rebuild to generate such new sequence.
2.  Download  the  file   with  PBD\_ID:  1AOI,  that  is,  chromatins
nucleosome core particle. Run it through \textrm{find\_pair}



\bibliography{biblio}




